The effects of context inference on motor learning: savings, de-adaptation and spontaneous recovery

### 2021-10-20 adaptation/article/article.tex
Removed redundancy. 1.5k words to go

\section{Introduction}
Humans are able to adapt their movements to changing dynamics of the environment and of the body itself. We adapt to slow changes in our body dynamics, e.g. because of muscle exhaustion throughout the day, but we can also adapt to sudden changes in the environment. For example, when picking up a heavy object, we must adapt our motor commands to be able to move the extra weight safely. Even before picking it up, the size and material of an object provide us with contextual information which we can use to prepare the correct posture for moving the object. However, contextual information cannot always be relied on, e.g. a large parcel with unknown content.

It has been shown that humans can learn multiple tasks during the course of a single experiment, as well as learn to adapt to often-opposing modifications to the dynamics of movements \citep{Gandolfo_Motor_1996,Shadmehr_Functional_1997}, although interference between tasks can slow down learning \citep{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}. Additionally, humans have been shown to dynamically switch between different learned adaptation \citep{Davidson_Scaling_2004,Ethier_Spontaneous_2008,Lee_Dual_2009}. The mechanisms behind these switches, the specifics of how these switches manifest in behavior, and the ways in which switching can be impeded or improved, are an ongoing, active topic of research.

@@

To explain these phenomena, a number of computational models have been introduced. The least complex of these models are linear learners \citep{Smith_Interacting_2006,Forano_Timescales_2020,Scheidt_Learning_2001}, in which a single one-dimensional variable is used to describe the adaptation of motor commands given the observed motor errors. By having multiple time scales, each with a learning rate and a forgetting rate, these models have been fitted to experimental data, partially explaining these distinct experimental phenomena.

Bayesian accounts of motor adaptation have also been presented, providing an alternative explanation for savings and quick de-adaptation in the form of switching between internal models of adaptation via predictions generated by these internal models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}. These models replace the idea of explicit multiple time scales with an estimation of their own uncertainty: as the the same task is practiced, the uncertainty on an estimate of the underlying modification is reduced, which lowers variability in responses and increases the stability of the estimate.

While these general models of adaptation explain the most common phenomena observed in experiments, other known phenomena remain outside of their scope. For example, it is known that adaptation rate is reduced in situations where the environment is unstable and unpredictable \citep{Herzfeld_memory_2014}, or situations in which errors are small \citep{Marko_Sensitivity_2012} and slowly introduced \citep{Huang_Persistence_2009}. Action selection has also been found to depend on the history of adaptations learned \citep{Vaswani_Decay_2013,Davidson_Scaling_2004}.

@@

To this end, we present a computational model for motor adaptation inspired by models such as MOSAIC \citep{Wolpert_Multiple_1998} and Bayesian switching models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}. We build on these models by expanding the concept of context and introducing an explicit mechanism for inferring the context based on Bayesian inference. In our model, the context encompasses all the information relevant to making decisions, both environmental and internal to the agent making these decisions. To infer the current context, our model integrates several sources of information, including the predictions made by its internal forward models, an estimate of the reliability of these predictions, direct sensory information (such as visual contextual cues), proprioceptive information, and feedback provided by the environment (e.g. whether the task was successfully completed). Having inferred the context, the model can choose the best-fitting internal model with which to make decisions.

@@

\subsubsection{Motor adaptation}
Motor adaptation refers to the ability of humans and other animals to adapt their motor commands based on performance errors. For example, we can learn to perform a reaching movement even when underwater, where the relatively high viscosity of water means that the motor commands we have learned all our lives no longer produce the desired effect.

It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals \citep{Wolpert_Multiple_1998}. Here, we use this principle and chose to make use of an exact Bayesian learner. As has been discussed before \citep{Oh_Minimizing_2019}, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also eliminating the need for multiple explicit time scales of learning as the learner can deal with adaptation after any number of trials. As we will show below this feature will be useful to explain specific key phenomena parsimoniously.

To describe this component of the model, we make use of the example of reaching movements. We define a forward model as follows:
\begin{equation}
p(\vartheta | s, c; \phi) = f(s, c; \phi) \label{eqn:forward-model}
\end{equation}
where $\vartheta$ are the predicted outcomes of an action (e.g. the final position of the hand in a reaching movement), $s$ is the current state (the current position of the hand), $c$ is a motor command and $\phi$ are the parameters of the forward model $f(\cdot)$. We assume that during learning the parameters $\phi$ have been fine tuned to produce an accurate forward model $f$ to guide our movements and it is these parameters which are updated during motor adaptation.

When an observed outcome does not match the predictions produced by $f$, the prediction error is used to update the parameters $\phi$ of the forward model. To describe this adaptation process, our model makes uses of Bayesian inference, such that
\begin{equation}
q(\phi | \vartheta, s, c) \propto p(\vartheta | s, c, \phi)p(\phi) \label{eqn:update-mini}
\end{equation}
where $q(\phi | ...)$ represents the new estimated probability distribution over the possible values of $\phi$ after having observed the outcome $\vartheta$ of the previous motor command $c$. The likelihood $p(\vartheta | s, c, \phi)$ is the probability of observing the outcome $\vartheta$ as predicted by the forward model $f$, and $p(\phi)$ is the previous estimate of $\phi$.

@@

Following these ideas, our model performs context inference that integrates several sources of information, including sensory cues, proprioceptive information, outcome prediction errors and reward prediction errors.

In the model, context is a categorical variable $\zeta$ which is inferred using Bayes' theorem:
\begin{equation}
q(\zeta | \vartheta_t, \vartheta_{t-1}, c_{t-1}) \propto p(\vartheta_t | \vartheta_{t-1}, c_{t-1})p(\zeta)
\label{eqn:context-inference-mini}
\end{equation}
where the prior distribution over contexts $p(\zeta)$ includes information from two sources: (1) an expectation of continuity, encoding the expectation that the context does not change from trial to trial, and (2) an overall expectation of observing any one context, if the context did change. Below, we discuss how these two sources of information effect different experimentally observed phenomena.

In the previous section, we used the variable $\vartheta$ to refer to the outcomes of actions. In this section, we expand the definition of an observation $\vartheta$ to include any information given to the decision-making agent by the environment. This includes, as before, the outcomes of actions (e.g. the new position of the hand), but also any other sensory information that might be indirectly related to actions, or even completely independent from them, as is the case for visual cues in experiments \cite[e.g.][]{Addou_Colored_2011}. For this section, the most important component of $\vartheta$ is contextual sensory information, i.e. information that might directly help infer the identity of the current context.

In experimental settings, the contextual sensory information can take the form of a visual cue \citep[e.g.][]{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \citep[e.g.][]{Forano_Timescales_2020,Shadmehr_Adaptive_1994} or even the start of a new block of trials \citep{Ethier_Spontaneous_2008}. All these sources of information, alongside the outcomes of previous actions and the priors $p(\vartheta)$ form part of $p(\vartheta_t | \vartheta_{t-1}, c_{t-1})$ and are integrated together to infer the identity of a context.

\subsubsection{Action selection}
Action selection is affected by the current context via the context-dependent forward model that is used. In our model, action selection is made by building a distribution over available actions which is a weighted sum of the distributions given by the existing forward models, where the weights are obtained by the context inference component of the model. In other words, a distribution is created as follows:
\begin{equation}
p(c_t) \propto \displaystyle\sum_{\zeta \in \Phi}q(\zeta | \vartheta_t, \vartheta_{t-1}, c_{t-1}) p(c_t | \zeta) \label{eq:action-selection}
\end{equation}
where $p(c_t | \zeta)$ corresponds to \eref{eqn:forward-model}. From this distribution, the current action (motor command $c_t$) can be sampled and carried out.

\subsubsection{From observation to action selection}
Each trial starts with a new observation $\vartheta_t$, which comprises contextual sensory information (when available), as well as the outcome of the previously-selected action.

The first step in the model is to perform context inference using the current observation $\vartheta_t$ according to \eref{eqn:context-inference-mini} and \eref{eqn:estimated-context}. Once a probability distribution over contexts has been created, all models are updated according to \eref{eqn:update-mini} and \eref{eqn:update-full}, based on the distribution over contexts. Finally, action selection is performed using the distribution over contexts as well as the updated forward models, following \eref{eq:action-selection}.

@@

In this section, we discuss savings, and the related phenomenon of de-adaptation, in terms of switching between forward models. This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019,Imamizu_Explicit_2007} where we show in addition that expanding forward model switching into context inference explains facets of savings that these previous models do not.

To show this, we examined multiple experimental studies in which savings are observed. We categorized these studies based on the amount of contextual information made available to participants into cued-context and partially-cued experiments: In some experiments \citep[e.g.][]{Kim_Neural_2015,Lee_Dual_2009}, the context is clearly revealed to the participant. We call these cued-context experiments. In other experiments, partial information is available to participants \citep[e.g.][]{Davidson_Scaling_2004,Zarahn_Explaining_2008} in the form of large prediction errors, partial contextual information or reward prediction errors; we refer to these as partially-cued experiments.

Because context inference integrates information from different sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use to infer the context. For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone. In fact, the mere act of holding the handle informs participants that something is different from baseline. Proprioceptive signals can also provide contextual information \cite{Dizio_Motor_1995,Shadmehr_Adaptive_1994}. The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than the expected trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \cite{Ethier_Spontaneous_2008}.

@@

We simulated data with our model fitting the experimental structure in \cite{Davidson_Scaling_2004}. We show the results in \fref{fig:davidson-2004}, alongside the experimental results from \cite{Davidson_Scaling_2004}. It can be seen that the agent exposed to the -A context shifts back to A more slowly than the one exposed to 3A, qualitatively reproducing the data from \cite{Davidson_Scaling_2004}, shown in \fref{fig:davidson-2004}B. Note that in our simulations, the difference between the groups, as well as the variability within each group, is smaller than in the experimental data. However, the effect can be reproduced with any number of simulated runs (each one representing one participant), pointing towards a reliable effect. While we focused here on a specific, well-known experiment, we discuss experiments similar to that by \cite{Davidson_Scaling_2004} in the Discussion.

@@

Further evidence for this effect of uncertainty about contexts can be found during error-clamp blocks in several experiments. We discuss these experiments in the following section. Additionally, we discuss experiments similar to that by \cite{Davidson_Scaling_2004} in the Discussion.

@@

In \fref{fig:vaswani-2013}A, we show data simulated with our model, following the parameters of the experiment by \cite{Vaswani_Decay_2013}, and in \fref{fig:vaswani-2013}B we show the experimental plots adapted from \cite{Vaswani_Decay_2013}. The displayed adaptation is shown during the E trials for the three experimental groups in the experiment. It can be seen that group 1.3 (i.e. the participants who had learned in the -A/2 context in addition to A) more quickly recognizes a change in context and lowers the force applied on the mechanical handle, as can be seen in the experimental data.

@@

Similar models have been presented before \cite[e.g.][]{Wolpert_Multiple_1998,Baddeley_System_2003,Imamizu_Explicit_2007}, which infer the internal models (forward and backward) that best explain the observations and use them to issue motor command. We built on these studies in two important ways: (1) we generalized the inference over the best-fitting internal models to the concept of context inference, which includes all sensory information, in addition to feedback provided by the experimenters or the task; (2) we showed how this seemingly small generalization provides a parsimonious explanation for several experimental phenomena across many experiments in motor adaptation that, up to now, had required multiple \textit{ad hoc} mechanisms.

@@

It is important to note that the most essential part of the model we presented is the concept of the context and the inclusion of a Bayesian inference scheme over the context, which integrates all the information that is available to the participant during the experiment. Most notably, we chose to make use of exact Bayesian inference for motor adaptation, with conjugate priors to derive exact update equations. This choice greatly simplified our simulations and provides intuitive interpretations of experimental results. However, we expect the results we presented here to hold using alternative models for motor adaptation, such as Kalman filter-based learners \cite[e.g.][]{Oh_Minimizing_2019,Baddeley_System_2003} and other Bayesian implementations \cite[e.g.][]{Wolpert_Multiple_1998,Kording_Bayesian_2004}. A well established family of models for motor adaptation is linear learners \cite[e.g.][]{Smith_Interacting_2006,Forano_Timescales_2020,Lee_Dual_2009}, however, these models do not incorporate uncertainty in the estimate of the parameters of the internal models, which would limit their implementation in our hierarchical model.

In what follows, we review the experimental literature to highlight the importance of contextual information and its effects on observed behavior during motor learning tasks, as shown by our model. In addition, we discuss several predictions that our model makes, which could be verified experimentally in future studies.

@@

The paradigm by \cite{Herzfeld_memory_2014} was designed to test the hypothesis that a memory of errors affected the speed with which participants adapted to curl forces. While it has been shown that sensitivity to error can be modulated \citep{Marko_Sensitivity_2012}, and indeed in our model it depends on the amount of experience (see \sref{sec:interpreting-hyperparameters}), to our knowledge there are no other experiments suggesting that it depends on the history of errors. As before, the introduction of clear sensory contextual cues would allow us to gauge the importance of context inference and of previous errors in the observed data. The same applies to the experiments by \cite{Davidson_Scaling_2004}.

\subsection{Further experimental evidence}
Context is often directly observable, either via environmental cues or via feedback observed after actions. In many behavioral experiments, for example, the context of the current block of trials is easily identifiable by the color of a cue light \citep[e.g.][]{Ethier_Spontaneous_2008}, the posture of the hand \citep[e.g.][]{Gandolfo_Motor_1996}, the target of the movement \citep[e.g.][]{Lee_Dual_2009} or by the workspace in which the task is being performed \citep[e.g.][]{Shadmehr_Adaptive_1994}. The context can also be selected by the agent itself. For example, a person can decide the level of care devoted to moving an egg or a lead ball around, which determines the way in which the person would move their hands. In these cases, the context is clearly determined and known to the agent and switches between different adaptations are immediate.

In other cases, however, the context might not be directly observable, and context inference takes the form of an evidence-accumulating process that can take any amount of time to be sure of the context. It is in these cases where the effects of context inference are most noticeable. While many experiments exist that give probabilistic contextual information \cite[e.g.][]{Scholz_uncontrolled_1999,Behrens_Learning_2007,Nassar_Dissociable_2019}, evidence accumulation is not limited to these explicit cases. Indeed, as we noted in the Results section, many experiments inadvertently include partial contextual information used by participants.

@@

In our account, if all information indicating a change in context is removed from the experiment, participants would continue to behave as they were in the previous block. Evidence for this can be seen in experiments 2 and 3 by \cite{Vaswani_Decay_2013}, where participants were shown random errors during E trials, with a variance matching that of previously observed motor commands. The authors showed that by matching the errors expected by participants, they eliminated the slow tapering off observed in most E blocks. As we showed in \fref{fig:vaswani-2013}D, this effect need not be all or none: in some experimental sessions, participants might fail to notice that a change has occurred due to the stochasticity of their own actions. This would lead to an observed delay before a participant changes her behavior after the E block has started.

@@

In addition to providing a unifying explanation for phenomena such as switching, savings, spontaneous recovery and slow return to baseline, our model makes unique predictions that could be tested in future experiments.

@@

Context inference can work through context-specific (possibly unreliable) sensory cues,
or through the dynamics of the system (e.g. expected outputs of motor
commands), using prediction error as the force behind inference. In this case,
the generative model can be written as follows:
\[
p(z_t, \vartheta_t | z_{t-1}, c_{t-1}) = p(\vartheta_t | z_t)p(z_t|\z_{t-1}, c_{t-1})
\]
and inferring the current (hidden) generalized state is done with Bayes' theorem:
\[
  q(z_t | \vartheta_t, \vartheta_{t-1}, c_{t-1}) \propto p(\vartheta_t | z_t)p(z_t|\vartheta_{t-1}, c_{t-1})
\]

Additionally, after having inferred the current state, motor adaptation is given by
\[
q(\phi | z_t, z_{t-1}, c_{t-1}) \propto p(z_t | z_{t-1}, c_{t-1}, \phi)q(z_{t-1}|\ldots)p(\phi)
\]
where $\phi$ are the parameters of the internal model for the dynamics, also called the forward model,
i.e. the function $f: (z_t, c_t) \rightarrow z_{t+1}$, which determines the beliefs of
the agent regarding how the system evolves after having taken action $c_t$. $q(z_{t-1}|\ldots)$ is the
posterior probability calculated in the previous trial.

@@

To complete the action-selection picture, we return to context inference. As
discussed above, action selection is made via a weighted sum of the motor commands
produced by each forward model:
\begin{equation}
p(c_t | ...) \propto \displaystyle \sum_{i} q(\zeta_i)p(c_t | \zeta_i, ...)
\end{equation}
From this distribution, an action is sampled at each trial.

>>>>

\section{Introduction}
It has been shown that humans can learn multiple tasks during the course of a single experiment, as well as learn to adapt to often-opposing modifications to the dynamics of movements \citep{Gandolfo_Motor_1996,Shadmehr_Functional_1997}, although interference between tasks can slow down learning \citep{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}. Additionally, humans have been shown to dynamically switch between different learned adaptation \citep{Davidson_Scaling_2004,Ethier_Spontaneous_2008,Lee_Dual_2009}.

@@

To explain these phenomena, a number of computational models have been introduced. The least complex of these models are linear learners \citep{Smith_Interacting_2006,Forano_Timescales_2020,Scheidt_Learning_2001}. Bayesian accountshave also been presented, providing an alternative explanation for savings and quick de-adaptation in the form of switching between internal models of adaptation via predictions generated by these internal models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.

While these general models of adaptation explain the most common phenomena observed in experiments, other known phenomena remain outside of their scope. For example, it is known that adaptation rate is reduced in situations where the environment is unstable and unpredictable \citep{Herzfeld_memory_2014}, or situations in which errors are small \citep{Marko_Sensitivity_2012} or adaptations slowly introduced \citep{Huang_Persistence_2009}. Action selection has also been found to depend on the history of adaptations learned \citep{Vaswani_Decay_2013,Davidson_Scaling_2004}.

@@

To this end, we present a computational model for motor adaptation inspired by models such as MOSAIC \citep{Wolpert_Multiple_1998} and Bayesian switching models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}. We build on these models by expanding the concept of context and introducing an explicit mechanism for inferring the context based on Bayesian inference. In our model, the context encompasses all the information relevant to making decisions, both environmental and internal to the agent making these decisions. To infer the current context, our model integrates several sources of information, including the predictions made by its internal forward models, an estimate of the reliability of these predictions, direct sensory information (such as visual contextual cues), proprioceptive information, and feedback provided by the environment (e.g. whether the task was successfully completed).

@@

\subsubsection{Motor adaptation}
Motor adaptation refers to the ability of humans and other animals to adapt their motor commands based on performance errors. For example, we can learn to perform a reaching movement when underwater, where the relatively high viscosity of water means that the motor commands we have learned all our lives no longer produce the desired effect.

It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals \citep{Wolpert_Multiple_1998}. Here, we use this principle and chose to make use of an exact Bayesian learner. To adapt to altered dynamics (e.g. a mechanical arm exerting a force on a participant's hand), our model updates the parameters $\phi$ (see Methods) of a context-specific forward model. These updates are done according to \eref{eqn:motor-adaptation}, using the difference between predictions made with the current parameters of the forward model, and the current observation. With this, a posterior estimate on $\phi$, $q(\phi)$, is calculated after each trial.

@@

Following these ideas, our model performs context inference that integrates several sources of information, including sensory cues, proprioceptive information, outcome prediction errors and reward prediction errors. As with motor adaptation, the context is inferred with exact Bayesian inference, with the updates given by \eref{eqn:estimated-context}.

This component of the model can be used to include contextual sensory information from many sources. In experimental settings, for example, the contextual sensory information can take the form of a visual cue \citep[e.g.][]{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \citep[e.g.][]{Forano_Timescales_2020,Shadmehr_Adaptive_1994} or even the start of a new block of trials \citep{Ethier_Spontaneous_2008}. All these sources of information, alongside the outcomes of previous actions and the prior beliefs over context $p(\vartheta)$ are integrated to infer the identity of a context.

\subsubsection{Action selection}
Action selection is affected by the current context via the context-dependent forward model that is used. In our model, action selection is made by building a distribution over available actions which is a weighted sum of the distributions given by the existing forward models, where the weights are obtained by the context inference component of the model.

In other words, a distribution $p(c_t)$ over all possible motor commands is created following \eref{eqn:dist-comm}, from which the next action (motor command $c_t$) can be sampled and carried out.

@@

In this section, we discuss savings, and the related phenomenon of de-adaptation, in terms of switching between forward models. This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019,Imamizu_Explicit_2007}. Here we show in addition that expanding forward model switching into context inference explains facets of savings that these previous studies did not.

To show this, we examined multiple experimental studies in which savings are observed. We categorized these studies based on the amount of contextual information made available to participants: In some experiments \citep[e.g.][]{Kim_Neural_2015,Lee_Dual_2009}, the context is clearly revealed to the participant. We call these cued-context experiments. In other experiments, partial information is available to participants \citep[e.g.][]{Davidson_Scaling_2004,Zarahn_Explaining_2008} in the form of large prediction errors, partial contextual information or reward prediction errors; we refer to these as partially-cued experiments.

Because context inference integrates information from different sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use to infer the context. For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone. In fact, the mere act of holding the handle informs participants that something is different from baseline. Proprioceptive signals can also provide contextual information \citep{Dizio_Motor_1995,Shadmehr_Adaptive_1994}. The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than the expected trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \cite{Ethier_Spontaneous_2008}.

@@

We simulated data with our model fitting the experimental structure in \cite{Davidson_Scaling_2004}. We show the results in \fref{fig:davidson-2004}, alongside the experimental results from \cite{Davidson_Scaling_2004}. It can be seen that the agent exposed to the -A context shifts back to A more slowly than the one exposed to 3A, qualitatively reproducing the data from \cite{Davidson_Scaling_2004}, shown in \fref{fig:davidson-2004}B. Note that in our simulations, the difference between the groups, as well as the variability within each group, is smaller than in the experimental data. However, the effect can be reproduced with any number of simulated runs (each one representing one participant), pointing towards a reliable effect.

@@

@@

In \fref{fig:vaswani-2013}A, we show data simulated with our model, following the parameters of the experiment by \cite{Vaswani_Decay_2013}, and in \fref{fig:vaswani-2013}B we show the experimental plots adapted from \cite{Vaswani_Decay_2013}. The displayed adaptation is shown during the E trials for the three experimental groups in the experiment. It can be seen that group 1.3 (i.e. the participants who had learned in the -A/2 context in addition to A) more quickly recognized a change in context and lowers the force applied on the mechanical handle, as can be seen in the experimental data.

@@

We generalized the inference over the best-fitting internal models to the concept of context inference, which includes all sensory information, in addition to feedback provided by the experimenters or the task. With this, we showed how this seemingly small generalization provides a parsimonious explanation for several experimental phenomena across many experiments in motor adaptation that, up to now, had required multiple \textit{ad hoc} mechanisms.

@@

Our choice of an exact Bayesian learner greatly simplified our simulations, but we expect the results we presented here to hold using alternative models for motor adaptation, such as Kalman filter-based learners \cite[e.g.][]{Oh_Minimizing_2019,Baddeley_System_2003} and other Bayesian implementations \cite[e.g.][]{Wolpert_Multiple_1998,Kording_Bayesian_2004}. A well established family of models for motor adaptation is linear learners \cite[e.g.][]{Smith_Interacting_2006,Forano_Timescales_2020,Lee_Dual_2009}, however, these models do not incorporate uncertainty in the estimate of the parameters of the internal models, which would limit their implementation in our hierarchical model.

@@

The paradigm by \cite{Herzfeld_memory_2014} was designed to test the hypothesis that a memory of errors affected the speed with which participants adapted to curl forces. While it has been shown that sensitivity to error can be modulated \citep{Marko_Sensitivity_2012}, and indeed in our model it depends on the amount of experience, to our knowledge there are no other experiments suggesting that it depends on the history of errors. As before, the introduction of clear sensory contextual cues would allow us to gauge the importance of context inference and of previous errors in the observed data. The same applies to the experiments by \cite{Davidson_Scaling_2004}.

\subsection{Further experimental evidence}
Context is often directly observable, either via environmental cues or via feedback observed after actions. It can also be selected by the agent itself. For example, a person can decide the level of care devoted to moving an egg or a lead ball around. In many other cases, however, context inference takes the form of an evidence-accumulating process that can take any amount of time to be sure of the context. It is in these cases where the effects of context inference are most noticeable. While many experiments exist that give probabilistic contextual information \cite[e.g.][]{Scholz_uncontrolled_1999,Behrens_Learning_2007,Nassar_Dissociable_2019}, evidence accumulation is not limited to these explicit cases. Indeed, as we noted in the Results section, many experiments inadvertently include partial contextual information used by participants.

@@

In our account, if all information indicating a change in context is removed from the experiment, participants would continue to behave as they were in the previous block. Evidence for this can be seen in experiments 2 and 3 by \cite{Vaswani_Decay_2013}, where participants were shown random errors during E trials, with a variance matching that of previously observed motor commands. The authors showed that by matching the errors expected by participants, they eliminated the slow tapering off observed in most E blocks.

@@

@@

Context inference can work through context-specific (possibly unreliable) sensory cues,
or through the dynamics of the system (e.g. expected outputs of motor
commands), using prediction error as the force behind inference. In this case,
the generative model can be written as follows:
\[
p(z_t, \vartheta_t | z_{t-1}, c_{t-1}) = p(\vartheta_t | z_t)p(z_t|z_{t-1}, c_{t-1})
\]
and inferring the current (hidden) generalized state is done with Bayes' theorem:
\[
  q(z_t | \vartheta_t, \vartheta_{t-1}, c_{t-1}) \propto p(\vartheta_t | z_t)p(z_t|\vartheta_{t-1}, c_{t-1})
\]

Additionally, after having inferred the current state, motor adaptation is given by
\begin{equation}
q(\phi | z_t, z_{t-1}, c_{t-1}) \propto p(z_t | z_{t-1}, c_{t-1}, \phi)q(z_{t-1}|\ldots)p(\phi) \label{eqn:motor-adaptation}
\end{equation}
where $\phi$ are the parameters of the internal model for the dynamics, also called the forward model,
i.e. the function $f: (z_t, c_t) \rightarrow z_{t+1}$, which determines the beliefs of
the agent regarding how the system evolves after having taken action $c_t$. $q(z_{t-1}|\ldots)$ is the
posterior probability calculated in the previous trial.

@@

To complete the action-selection picture, we return to context inference. As
discussed above, action selection is made via a weighted sum of the motor commands
produced by each forward model:
\begin{equation}
p(c_t | ...) \propto \displaystyle \sum_{i} q(\zeta_i)p(c_t | \zeta_i, ...) \label{eqn:dist-comm}
\end{equation}
From this distribution, an action is sampled at each trial.
